---
title: "confidence_pm_analysis"
author: "Alan Voodla, Andero Uusberg"
date: "30 10 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

Title of the article

The analysis is split into the following parts

1. Behavioral level analyses
2. The effects of task manipulations on PM-ERPs
3. PM-ERPs and confidence correlation analyses
4. Mediation analyses


```{r}
# Load libraries and data (2 separate ERP dataset exports - 1. for ERN & Pe (d1), 2. for P3 (d2)   )
library(tidyverse)
library(readxl)
library(bmlm)
library(lme4)
library(apaTables)
library(jtools)
library(ez)
library(lmerTest)
library(ggpubr)
library(rmcorr)
library(bayesplot)

# import data
# ERN & pe dataset

d1 <- read_csv("data/04-Nov-2020-ERN_PE_export.csv")
  
# P3 dataset
d2 <- read_csv("data/04-Nov-2020-P3export.csv")

# remove variables with NA-s in confidence

table(is.na(d1$confidence))

table(is.na(d2$confidence))

d1 = d1 %>% na.omit()
d2 = d2 %>% na.omit()

table(is.na(d1$confidence))
table(is.na(d2$confidence))

# recode variables for R

d1$participant = as.factor(d1$participant)
levels(d1$participant)

d2$participant = as.factor(d2$participant)
levels(d2$participant)

d1$response = as.factor(d1$response)
d2$response = as.factor(d2$response)

# summary of confidence for both datasets

summary(d1$confidence)
summary(d2$confidence)

# participants
d1 %>% group_by(Subject) %>% summarize(n())
d2 %>% group_by(Subject) %>% summarize(n())

```

# Analysis of trials per condition and participant\

```{r}
trialtable = d1 %>% group_by(participant, difficulty, fluency) %>%
  dplyr::summarize(trialcount = n())

summary(trialtable$trialcount)

trialtable2 = d2 %>% group_by(participant, difficulty, fluency) %>%
  dplyr::summarize(trialcount = n())

summary(trialtable2$trialcount)



# Manipulation effects on accuracy, rt , confidence

# logistic regression mixed model regression for accuracy


#fit_acc = glmer(response ~ fluency*difficulty + (fluency:difficulty | participant), family=binomial, data=d1) 
#save(fit_acc, file="analysis/fit_acc.rData")
load("analysis/fit_acc.rData")
summary(fit_acc)

# accuracy effect

table_sum <- d1 %>% group_by(participant, difficulty, fluency) %>%
summarise("Mean_response_time"= mean(response_time, na.tm=T),
          "Mean_confidence" = mean(confidence, na.rm=T), 
          "Mean_accuracy" = round(sum(response=="correct")/length(response) * 100,1))

# 95% ci helper function for plot
mean_ci = function(x) {
	ci = qt(0.975,df=length(x)-1)*sd(x)/sqrt(length(x))
	return(data.frame(y = mean(x), ymin = mean(x) - ci, ymax = mean(x) + ci))
}

plot1 = ggplot(table_sum, aes(y = Mean_accuracy, x = difficulty, fill=fluency, color = fluency)) +
  stat_summary(fun.data = "mean_ci", position = position_dodge(0.1), geom = "errorbar", size = 1, width = 0.2, alpha = 0.3) +
  stat_summary(fun.y = mean, geom = "point",position = position_dodge(0.1), size = 2, alpha = 0.8) +
  scale_y_continuous(limits = c(0,100)) +
  theme_apa(legend.use.title = TRUE, legend.pos = "topright") +
  theme(text=element_text(size=12), axis.title.x = element_text(size = 12, face = "bold"),
  axis.title.y = element_text(size = 12, face = "bold")) +
  ylab("Mean response accuracy in %") +
  xlab("Trial difficulty") +
  labs(fill = "Fluency", color = "Fluency")

plot1

# Effects of manipulations on RTs

fit_rt = ezANOVA(data = table_sum,
													dv = .(Mean_response_time), 
													wid = .(participant),
													within = .(fluency,difficulty),
													detailed = TRUE,
                          return_aov = T)
apa.ezANOVA.table(fit_rt)


plot2 = ggplot(table_sum, aes(y = Mean_response_time, x = difficulty, fill=fluency, color = fluency)) +
  stat_summary(fun.data = "mean_ci", position = position_dodge(0.1), geom = "errorbar", size = 1, width = 0.2, alpha = 0.3) +
  stat_summary(fun.y = mean, geom = "point",position = position_dodge(0.1), size = 2, alpha = 0.8) +
  scale_y_continuous(limits = c(0,2)) +
  theme_apa(legend.use.title = TRUE, legend.pos = "bottomright", legend.font.size = 12) +
  theme(text=element_text(size=12), axis.title.x = element_text(size = 12, face = "bold"),
  axis.title.y = element_text(size = 12, face = "bold")) +
  ylab("Mean response time (s)") +
  xlab("Trial difficulty") + 
  labs(fill = "Fluency", color = "Fluency")

plot2



## manipulation effects on confidence ### 

fit2 = ezANOVA(data = table_sum,
													dv = .(Mean_confidence), 
													wid = .(participant),
													within = .(fluency,difficulty),
													detailed = TRUE,
                          return_aov = T)


apa.ezANOVA.table(fit2)

# mixed model 

fit_confidence = lmer(Mean_confidence ~ difficulty*fluency + (1|participant), data=table_sum)


summary(fit_confidence)


# plot3 

plot3 = ggplot(table_sum, aes(y = Mean_confidence, x = difficulty, fill=fluency, color = fluency)) +
  stat_summary(fun.data = "mean_ci", position = position_dodge(0.1), geom = "errorbar", size = 1, width = 0.2, alpha = 0.3) +
  stat_summary(fun.y = mean, geom = "point",position = position_dodge(0.1), size = 2, alpha = 0.8) +
  scale_y_continuous(limits = c(0,100)) +
  theme_apa(legend.use.title = TRUE, legend.pos = "topright", legend.font.size = 12) +
  theme(text=element_text(size=12), axis.title.x = element_text(size = 12, face = "bold"),
  axis.title.y = element_text(size = 12, face = "bold")) +
  ylab("Mean confidence rating") +
  xlab("Trial difficulty") +
  labs(fill = "Fluency", color = "Fluency")

plot3

## combine plots for article ## 

#fig2 = ggarrange(plot1, plot2, plot3, ncol = 3,nrow = 1, labels = c("A", "B", "C") ,common.legend #= FALSE)
#fig2
### EXPORT ####

#tiff("new_figure.tiff", units = "cm", width=20, height=15, res=300)
#fig2
#dev.off()


```


Mixed effect models - effects of task manipulations on Pe, ERN and P3

```{r}


# effect of manipulations on ERN 
fit_ern <- lmer(ERN_FzCz ~ difficulty * fluency + (1|participant), data=d1)
summary(fit_ern)

## CI function
mean_ci = function(x) {
	ci = qt(0.975,df=length(x)-1)*sd(x)/sqrt(length(x))
	return(data.frame(y = mean(x), ymin = mean(x) - ci, ymax = mean(x) + ci))
}

# interaction plot       
ern_plot = ggplot(d1, aes(y = ERN_FzCz, x = as.factor(difficulty), fill= as.factor(fluency), color = as.factor(fluency))) +
  stat_summary(fun.data = "mean_ci", position = position_dodge(0.1), geom = "errorbar", size = 1, width = 0.2, alpha = 0.3) +
  scale_y_continuous() +
  stat_summary(fun.y = mean, geom = "point",position = position_dodge(0.1), size = 2, alpha = 0.8) +
  theme_apa(legend.use.title = TRUE, legend.pos = "topright") +
  theme(text=element_text(size=12), axis.title.x = element_text(size = 12, face = "bold"),
  axis.title.y = element_text(size = 12, face = "bold")) +
  ylab("Mean pooled ERN (µV) ") +
  xlab("Trial difficulty") +
  labs(fill = "Fluency", color = "Fluency")
ern_plot

######

# Pe


fit_Pe <- lmer(Pe_PzCz ~ difficulty * fluency + (1|participant), data=d1)
summary(fit_Pe)


# plot2 


pe_plot = ggplot(d1, aes(y = Pe_PzCz, x = as.factor(difficulty), fill= as.factor(fluency), color = as.factor(fluency))) +
  stat_summary(fun.data = "mean_ci", position = position_dodge(0.1), geom = "errorbar", size = 1, width = 0.2, alpha = 0.3) +
  scale_y_continuous() +
  stat_summary(fun.y = mean, geom = "point",position = position_dodge(0.1), size = 2, alpha = 0.8) +
  theme_apa(legend.use.title = TRUE, legend.pos = "topright") +
  theme(text=element_text(size=12), axis.title.x = element_text(size = 12, face = "bold"),
  axis.title.y = element_text(size = 12, face = "bold")) +
  ylab("Mean pooled Pe (µV) ") +
  xlab("Trial difficulty") +
  labs(fill = "Fluency", color = "Fluency")
pe_plot


##### P3 ##############
fit_P3 <- lmer(P3_Pz ~ difficulty * fluency + (1|participant), data=d2)
summary(fit_P3)


# plot2 


p3_plot = ggplot(d2, aes(y = P3_Pz, x = as.factor(difficulty), fill= as.factor(fluency), color = as.factor(fluency))) +
  stat_summary(fun.data = "mean_ci", position = position_dodge(0.1), geom = "errorbar", size = 1, width = 0.2, alpha = 0.3) +
  scale_y_continuous() +
  stat_summary(fun.y = mean, geom = "point",position = position_dodge(0.1), size = 2, alpha = 0.8) +
  theme_apa(legend.use.title = TRUE, legend.pos = "topright") +
  theme(text=element_text(size=12), axis.title.x = element_text(size = 12, face = "bold"),
  axis.title.y = element_text(size = 12, face = "bold")) +
  ylab("Mean pooled P3 (µV) ") +
  xlab("Trial difficulty") +
  labs(fill = "Fluency", color = "Fluency")
p3_plot




  
# combine plots for export

#library(ggpubr)
#fig3 = ggarrange(plot1, plot2, ncol = 2,nrow = 1, labels = c("A", "B") ,common.legend = TRUE)
#png("new_figure.png", units = "cm", width=20, height=15, res=300)
#fig3
#dev.off()



```







Correlation analysis for confidence and ERPs

```{r}
options(scipen=9999)
# ern & confidence (0.05, significant)
rmc1 <- rmcorr(participant = participant, measure1 = confidence, measure2 = ERN_FzCz , dataset = d1)
rmc1
plot(rmc1, d, overall = F, lty = 2, xlab = "Confidence", ylab = "ERN")

# PE & confidence (-0.02, non significant)
rmc2 <- rmcorr(participant = participant, measure1 = confidence, measure2 = Pe_PzCz , dataset = d1)
rmc2
plot(rmc2, d, overall = F, lty = 2, xlab = "Confidence", ylab = "Pe")

# P3 & confidence (-0.006, nonsignificant)
rmc3 <- rmcorr(participant = participant, measure1 = confidence, measure2 = P3_Pz , dataset = d2)
rmc3
plot(rmc3, d, overall = F, lty = 2, xlab = "Confidence", ylab = "P3")

# rmcorss for 2 levels of difficulty

# easy trials RM correlation confidence and ERN (significant corr for easy trials)
d1 %>% filter(difficulty == "easy") %>% rmcorr(participant = participant, measure1 = confidence, measure2 = ERN_FzCz)

# hard trials RM correlation confidence and ERN (ns for hard trials)
d1 %>% filter(difficulty == "hard") %>% rmcorr(participant = participant, measure1 = confidence, measure2 = ERN_FzCz)

# easy trials RM correlation confidence and Pe (ns for easy trials)
d1 %>% filter(difficulty == "easy") %>% rmcorr(participant = participant, measure1 = confidence, measure2 = Pe_PzCz)

# hard trials RM correlation confidence and ERN 
d1 %>% filter(difficulty == "hard") %>% rmcorr(participant = participant, measure1 = confidence, measure2 = Pe_PzCz)

# easy trials RM correlation confidence and P3 
d2 %>% filter(difficulty == "easy") %>% rmcorr(participant = participant, measure1 = confidence, measure2 = P3_Pz)

# hard trials RM correlation confidence and P3 
d2 %>% filter(difficulty == "hard") %>% rmcorr(participant = participant, measure1 = confidence, measure2 = P3_Pz)


############# error vs correct trials correlations #############################
# error trials vs correct trials (ns corr)
d1 %>% filter(response == "error") %>% rmcorr(participant = participant, measure1 = confidence, measure2 = ERN_FzCz) 

# error trials & Pe (ns corr)
d1 %>% filter(response == "error") %>% rmcorr(participant = participant, measure1 = confidence, measure2 = Pe_PzCz)

# correct trials and ERN (correlation - 0.06, significant)

d1 %>% filter(response == "correct") %>% rmcorr(participant = participant, measure1 = confidence, measure2 = ERN_FzCz) 
# correct trias and Pe
d1 %>% filter(response == "correct") %>% rmcorr(participant = participant, measure1 = confidence, measure2 = Pe_PzCz)
#### P3 ####
# error trials & P3 
d2 %>% filter(response == "error") %>% rmcorr(participant = participant, measure1 = confidence, measure2 = P3_Pz)

# correct trials and P3

d2 %>% filter(response == "correct") %>% rmcorr(participant = participant, measure1 = confidence, measure2 = P3_Pz)

```


Mediation analyses

Does pooled ERN mediate the effect of difficulty on confidence?

```{r}
#  define dataframe
med_ERN <- isolate(d = d1,
        by = "participant",   
        value = c("ERN_FzCz","confidence"))

med_ERN$difficulty = as.numeric(as.factor(med_ERN$difficulty))

med_ERN$fluency = as.numeric(as.factor(med_ERN$fluency))
                               
                               
# fit 

fit1 <- mlm(d=med_ERN,
           id = "participant",          # subject variable
           x = "difficulty",            # experimental condition
           m = "ERN_FzCz_cw",                  # mediator
           y = "confidence_cw",
           iter = 10000,
           cores = 4)                             

# convergence check

#pars <- c("a", "b", "cp", "corrab")
#mcmc_trace(as.data.frame(fit), pars = pars)



# summary
mlm_summary(fit1)               

# path plot
mlm_path_plot(fit1, xlab = "Difficulty", mlab = "ERN", ylab = "Confidence")

# within subject parameters
mlm_pars_plot(fit, pars = "u_cp", type = "coef", level = .95)

```

Does pooled ERN mediate the effect of fluency on confidence?


```{r}

fit2 <- mlm(d=med_ERN,
           id = "participant",          # subject variable
           x = "fluency",            # experimental condition
           m = "ERN_FzCz_cw",                  # mediator
           y = "confidence_cw",
           iter = 10000,
           cores = 4)

pars <- c("a", "b", "cp", "corrab")
mcmc_trace(as.data.frame(fit2), pars = pars)

mlm_summary(fit2)

mlm_path_plot(fit2, xlab = "Fluency", mlab = "Pooled ERN", ylab = "Confidence")

mlm_pars_plot(fit2, pars = c("u_pme"), type = "coef", level = .95)

```


Does pooled ERN mediate the effect of fluency on confidence on easy trials?


```{r}

easy_ERN = med_ERN %>% filter(difficulty=="1")



fit3 <- mlm(d=easy_ERN,
           id = "participant",          # subject variable
           x = "fluency",            # experimental condition
           m = "ERN_FzCz_cw",                  # mediator
           y = "confidence_cw",
           iter = 10000,
           cores = 4)

pars <- c("a", "b", "cp", "corrab")

mlm_path_plot(fit3, xlab = "Fluency", mlab = "Pooled ERN", ylab = "Confidence")

mlm_pars_plot(fit3, pars = c("u_b"), type = "coef", level = .95)


```





Difficulty mediation for error & correct trials
```{r}

ERN_error = med_ERN %>% filter(response == "error")

                             
                              
# fit difficulty mediation model

fit4 <- mlm(d= ERN_error,
           id = "participant",          # subject variable
           x = "difficulty",            # experimental condition
           m = "ERN_FzCz_cw",                  # mediator
           y = "confidence_cw",
           iter = 10000,
           cores = 4)                             



# summary
mlm_summary(fit4)                              # dependent variable

# path plot
mlm_path_plot(fit4, xlab = "Difficulty", mlab = "ERN", ylab = "Confidence")


# difficulty & ERN mediation model for correct trials

# select only correct responses

ERN_correct = med_ERN %>% filter(response == "correct")

# fit difficulty mediation model

fit5 <- mlm(d=ERN_correct,
           id = "participant",          # subject variable
           x = "difficulty",            # experimental condition
           m = "ERN_FzCz_cw",                  # mediator
           y = "confidence_cw",
           iter = 10000,
           cores = 4)                             


# summary
mlm_summary(fit5)                              # dependent variable

# path plot
mlm_path_plot(fit5, xlab = "Difficulty", mlab = "ERN", ylab = "Confidence")


```

Fluency mediation model for error & correct
```{r}

# select only error responses

flu_error = med_ERN %>% filter(response == "error")

                               
                               
# fit fluency-error mediation model

fit_flu_error <- mlm(d=flu_error,
           id = "participant",          # subject variable
           x = "fluency",            # experimental condition
           m = "ERN_cw",                  # mediator
           y = "confidence_cw",
           iter = 10000,
           cores = 4)                             



# summary
mlm_summary(fit_flu_error)                              # dependent variable

# path plot
mlm_path_plot(fit_flu_error, xlab = "Fluency", mlab = "Pooled ERN", ylab = "Confidence")


# correct, fluency ERN

flu_correct = med_ERN %>% filter(response == "correct")

                               
# fit fluency mediation model for correct trials

fit_flu_correct <- mlm(d=flu_correct,
           id = "participant",          # subject variable
           x = "fluency",            # experimental condition
           m = "ERN_cw",                  # mediator
           y = "confidence_cw",
           iter = 10000,
           cores = 4)                             



# summary
mlm_summary(fit_flu_correct)                              # dependent variable

# path plot
mlm_path_plot(fit_flu_correct, xlab = "Fluency", mlab = "Pooled ERN", ylab = "Confidence")


```









